### 2022-01-07 draft.tex
Deploying to gh-pages from @ gottacatchenall/ms_neutral_landscapes_jl@0303cabf9d137dc149a9b232d9cffde6e3e86443 🚀

In

@@

@@

\textbf{Fig 2}: Benchmark comparison of selected methods in each of the
three languages

In fig 2 we should a selection of neutral landscape generators (random,
edge gradient, perlin noise, distance-gradient)

\hypertarget{mpd-comparison}{%
\subsection{MPD comparison}\label{mpd-comparison}}

Why use this particular generator as the comparison? It's slow. So slow
that NLMR implements it in C++. (NLMR implements both MPD, neighbor,
randrect, and random neighboorhood in c++). Still these three
algorithms, which cosinsts of 3/16 of NLMR's alg implementations,
constitute 33\% of its codebase.

In this section we show our implementation of MPD is faster than NLMR's
C++ MPD across all scales, up to the asyptotic limit imposed by the
\(O(n^2)\) scaling of the raster

\begin{figure}
\centering
\includegraphics{./figures/draftfig3.png}
\caption{Comparison of speed of generating a midpoint displacement
neutral landscape (y-axis) against raster size (measured as length of
the size of a square raster, x-axis)}
\end{figure}

@@

\(M_t = f(M_{t-1})\)

>>>>

\begin{longtable}[]{@{}llllll@{}}
\toprule
\begin{minipage}[b]{0.35\columnwidth}\raggedright
Model\strut
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\raggedright
nlmpy?\strut
\end{minipage} & \begin{minipage}[b]{0.05\columnwidth}\raggedright
NLMR\strut
\end{minipage} & \begin{minipage}[b]{0.33\columnwidth}\raggedright
Description\strut
\end{minipage} & \begin{minipage}[b]{0.02\columnwidth}\raggedright
Reference\strut
\end{minipage} & \begin{minipage}[b]{0.02\columnwidth}\raggedright
Aliases\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.35\columnwidth}\raggedright
No gradient\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.33\columnwidth}\raggedright
Each cell is drawn randomly\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Planar gradient\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.33\columnwidth}\raggedright
A gradient from low to high at a given angle\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Distance gradient\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.33\columnwidth}\raggedright
Each cell is the distance between that cell and a location\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Random rectuangular cluster\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.33\columnwidth}\raggedright
Covers the plane in random rectanges until covered\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Random element nearest-neighbor\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedright
x*\strut
\end{minipage} & \begin{minipage}[t]{0.33\columnwidth}\raggedright
Discrete categories based on distance a set of \texttt{n} random
points\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\texttt{nlm\_mosaictess}, k-means\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Random cluster nearest-neighbor\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.33\columnwidth}\raggedright
Starts with \(n\) seed points and grows clusters probabilistically\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Random curds\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.33\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Gaussian Field\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.33\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Diamond-square\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.33\columnwidth}\raggedright
Improvement on Diamond-square and fractal brownian motion.\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Perlin noise\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.33\columnwidth}\raggedright
Method for noise with ``smoother'' features than DS/MPD\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Mosaic random field\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.33\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

@@

\begin{verbatim}
using NeutralLandscapes
using SimpleSDMLayers

quebec = SimpleSDMPredictor(WorldClim, BioClim; left=-90., right=-50., top=75., bottom=40.)
qcmask = fill(true, size(quebec)) 
qcmask[findall(isnothing, quebec.grid)] .= false

pltsettings = (cbar=:none, frame=:box)

plot(
    heatmap(rand(MidpointDisplacement(0.8), size(layer), mask=qcmask); pltsettings),
    heatmap(rand(PlanarGradient(), size(layer), mask=qcmask); pltsettings),
    heatmap(rand(PerlinNoise((4,4)), size(layer), mask=qcmask); pltsettings),
    heatmap(rand(NearestNeighborCluster(0.5), size(layer), mask=qcmask); pltsettings),
    dpi=400
)

savefig("interoperable.png")
\end{verbatim}

\begin{figure}
\centering
\includegraphics{./figures/interoperable.png}
\caption{todo}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{./figures/benchmark.png}
\caption{todo}
\end{figure}

@@

\(M_t = M_{t-1} + f(M(t-1))\)

\hypertarget{models-of-change}{%
\subsection{Models of change}\label{models-of-change}}

\hypertarget{directional}{%
\subsubsection{Directional}\label{directional}}

\hypertarget{temporally-autocorrelation}{%
\subsubsection{Temporally
autocorrelation}\label{temporally-autocorrelation}}

\(r\): rate, \(v\): variability, \(U\) matrix of draws from standard
\(\text{Normal}(0,1)\)

\(f_{T}(M_{ij}) = r + vU_{ij}\)

\hypertarget{spatial-autocorrelation}{%
\subsubsection{Spatial autocorrelation}\label{spatial-autocorrelation}}

\(r\): rate, \(v\): variability, \([Z(\delta)]_{ij}\): the \((i,j)\)
entry of the zscore of the \(\delta\) matrix

\(f_{S}(M_{ij}) = r + v \cdot [Z(\delta)]_{ij}\)

\hypertarget{spatiotemporal-autocorrelation}{%
\subsubsection{Spatiotemporal
autocorrelation}\label{spatiotemporal-autocorrelation}}

\hypertarget{rescaling-to-mimic-real-data}{%
\subsection{Rescaling to mimic real
data}\label{rescaling-to-mimic-real-data}}
